---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eclust

<!-- badges: start -->
<!-- badges: end -->

The goal of eclust is to implement the clustering algorithm from [Davis, R.A., Fernandes, L. and Fokianos, K. (2023)](https://onlinelibrary.wiley.com/doi/10.1111/jtsa.12688).

## Installation

You can install the development version of eclust like so:

``` r
#install.packages("pak")
pak::pkg_install("leonfernandes/eclust")
```

## Demonstration for `gdp`

We demonstrate the clustering method for the `gdp` dataset provided with this package, sourced from [https://www.conference-board.org/us/](https://www.conference-board.org/us/).

```{r setup}
library(eclust)
library(tsibble)
gdp
```

We preprocess the data to obtain the log returns of the gdp.

```{r}
gdp_log_return <-
  gdp |>
  dplyr::group_by(country) |>
  dplyr::mutate(gdp = log(gdp) - dplyr::lag(log(gdp))) |>
  tidyr::drop_na() |>
  dplyr::mutate(gdp = as.numeric(scale(gdp))) |>
  dplyr::rename(gdp_logreturn = gdp) |>
  dplyr::ungroup()
```

Calculate the energy distance based dissimilarity matrix.

```{r}
dist_mat <- edist(gdp_log_return, lag = 1)
```

Perform hierarchical clustering and display the obtained dendrogram.

```{r}
library(ggplot2)
hc <- stats::hclust(dist_mat, method = "ward.D2")
plot(hc)
```